Class-based network segregation, Economic Inequality and Redistributive Preferences across societies

### 2024-03-29 paper1_jiturra.Rmd
update detalles

Figure \@ref(fig:fig1) depicts the distribution of redistributive preferences across countries and social classes. First, it is possible to notice that the United States has lower redistributive preferences, whereas Russia is a society where redistributive preferences are stronger among all social classes. As was expected, in most societies, the working class holds stronger redistributive preferences in contrast to the intermediate and services classes, with four exceptions where the differences are close to zero between both classes. However, it is also notable that there are some differences between the two extreme cases. For example, in the U.S., the working class held lower redistributive preferences (49.1) than the intermediate class (50.6), but still above the service class (46.8). In contrast, in Russia, the general pattern of higher preferences among the working class is maintained, with an escalated decrease in the working (85.5), intermediate (83.5), and service class (83.0). Another interesting fact is related to class differences in redistributive preferences. For example, the differences between the service and the working class are close to 2.3 points in the United States. At the same time, Finland represents the average case in the distribution but is also one of the most polarized societies in terms of class-based differences in redistributive preferences, showing a gap of 14.1 points between the upper and the lower classes. Interestingly, Russian society depicts class differences similar to the U.S., with an average difference of 2.4 points.

@@

Moving to the bivariate macro relationships, Figure \@ref(fig:fig3) depicts the correlation between network homogeneity, redistributive preferences, and income inequality, our main societal characteristic of interest. First, according to @cohen_statistical_1988, for effect size criteria, we observe a medium positive and significant association between network homogeneity and redistributive preferences in Panel A (_r_ = 0.45, _p_ <.001). In concordance with the previously depicted distribution, we know that higher levels of network homogeneity are driven by the highly segregated networks of the working classes; what makes it logical that in countries where network homogeneity is high, this might also reflect greater social isolation among the working classes and consequently drive higher redistributive preferences. Second, Panel B depicts a positive but relatively weak and nonsignificant association between income inequality and network homogeneity (_r_ = 0.28, _p_<.01). In other words, average homogeneity also tends to increase in societies with greater income gaps. However, the macro association only shows the general pattern between countries but does not allow us to look deeper into class differences according to the levels of inequality. Interestingly, certain countries with higher levels of inequality and network homogeneity are societies that present class differences between the working class and the services class (e.g., Thailand or Mexico). By contrast, more egalitarian countries tend to be less homogeneous and have smaller class differences (e.g., Denmark or Sweden). As an additional exploratory analysis, we have found that the association between income inequality and the gap in network homogeneity between the upper and lower classes tends to be moderate and statistically significant (_r_ = 0.29, _p_<.01). In other words, income inequality not only increase homogeneity overall but higher social distance between social classes as well.

@@

texreg(digclas1_macro,stars = c(0.001, 0.01, 0.05, 0.1),symbol = "+",scalebox = 0.7,
       caption = paste("(\\#tab:table2)","Multilevel models for Income Inequality, Network segregation and Redistributive Preferences"),
       custom.note = "Note: Models include sampling weights and individual level controls centered within cluster (group mean). Standard errors in parentheses. %stars",
       caption.above = T,leading.zero = T,single.row = T,
       custom.coef.map = ccoef,
      groups = list("Social Class (Ref.= Service Class)" = 2:3,"Macro-level factors"=4:6, "Homogeneity*Social Class"=7:8, "Homogeneity * Social Class * Income Inequality"=10:11),
       float.pos = "h!",
      custom.gof.rows = list("Controls"=rep("Yes",4)),include.loglik = FALSE,include.aic = FALSE,
      custom.gof.names = c(NA,NA,"Num. groups",
                           "Var: Group (Intercept)",
                           "Var: Group Homogeneity",
                           "Var: Group Intermediate Class ",
                           "Var: Group Working Class",
                           "Cov: Group (Intercept), Homogeneity",
                           "Cov: Group (Intercept), Intermediate Class ",
                           "Cov: Group (Intercept), Working Class",
                           "Cov: Group Homogeneity, Intermediate Class",
                           "Cov: Group Homogeneity, Working Class",
                           "Cov: Group Intermediate Class, Working Class",
                           "Var: Residual"),
       use.packages = F)
```

@@

st(dfreg,labels=T,
   title="Descriptive Statistics for Study Variables",
   out = "latex",
   digits = 1, 
   factor.counts = F,
   factor.numeric = F,
   summ = list(
     c('notNA(x)','mean(x)','sd(x)','min(x)','max(x)'),
     c('notNA(x)','mean(x)')
   ),
   summ.names = list(
     c('N','Mean','SD','Min','Max'),
     c('Count','Percent')   
   ))

@@

## Inequality Ratios

@@

## Alternative: Social Distance

@@

## OECD Countries

>>>>

Figure \@ref(fig:fig1) depicts the distribution of redistributive preferences across countries and social classes. First, it is possible to notice that the United States has lower redistributive preferences, whereas Russia is a society where redistributive preferences are stronger among all social classes. As was expected, in most societies, the working class holds stronger redistributive preferences in contrast to the intermediate and services classes, with four exceptions where the differences are close to zero between both classes. However, it is also notable that there are some differences between the two extreme cases. For example, in the U.S., the working class held lower redistributive preferences (49.1) than the intermediate class (50.6), but still above the service class (46.8). In contrast, in Russia, the general pattern of higher preferences among the working class is maintained, with an escalated decrease in the working (85.5), intermediate (83.5), and service class (83.0).

Another interesting fact is related to class differences in redistributive preferences. For example, the differences between the service and the working class are close to 2.3 points in the United States. At the same time, Finland represents the average case in the distribution but is also one of the most polarized societies in terms of class-based differences in redistributive preferences, showing a gap of 14.1 points between the upper and the lower classes. Interestingly, Russian society depicts class differences similar to the U.S., with an average difference of 2.4 points.

@@

Moving to the bivariate macro relationships, Figure \@ref(fig:fig3) depicts the correlation between network homogeneity, redistributive preferences, and income inequality, our main societal characteristic of interest. First, according to @cohen_statistical_1988, for effect size criteria, we observe a medium positive and significant association between network homogeneity and redistributive preferences in Panel A (_r_ = 0.45, _p_ <.001). In concordance with the previously depicted distribution, we know that higher levels of network homogeneity are driven by the highly segregated networks of the working classes; what makes it logical that in countries where network homogeneity is high, this might also reflect greater social isolation among the working classes and consequently drive higher redistributive preferences. Second, Panel B depicts a positive but relatively weak and nonsignificant association between income inequality and network homogeneity (_r_ = 0.28, _p_<.01). In other words, average homogeneity also tends to increase in societies with greater income gaps. However, the macro association only shows the general pattern between countries but does not allow us to look deeper into class differences according to the levels of inequality.

Interestingly, certain countries with higher levels of inequality and network homogeneity are societies that present class differences between the working class and the services class (e.g., Thailand or Mexico). By contrast, more egalitarian countries tend to be less homogeneous and have smaller class differences (e.g., Denmark or Sweden). As an additional exploratory analysis (see Fig. \@ref(fig:figapp1)), we have found that the association between income inequality and the gap in network homogeneity between the upper and lower classes tends to be moderate and statistically significant (_r_ = 0.29, _p_<.01). In other words, income inequality not only increase homogeneity overall but higher social distance between social classes as well.

@@

texreg(digclas1_macro,stars = c(0.001, 0.01, 0.05, 0.1),symbol = "+",scalebox = 0.7,
       caption = paste("(\\#tab:table2)","Multilevel models for Income Inequality, Network segregation and Redistributive Preferences"),
       custom.note = "Note: Models include sampling weights and individual level controls centered within cluster (group mean). Standard errors in parentheses. %stars",
       caption.above = T,leading.zero = T,single.row = T,
       custom.coef.map = ccoef,
      groups = list("Social Class (Ref.= Service Class)" = 2:3,"Macro-level factors"=4:6, "Homogeneity x Social Class"=7:8, "Homogeneity x Social Class x Income Inequality"=10:11),
       float.pos = "h!",
      custom.gof.rows = list("Controls"=rep("Yes",4)),include.loglik = FALSE,include.aic = FALSE,
      custom.gof.names = c(NA,NA,"Num. groups",
                           "Var: Country (Intercept)",
                           "Var: Country Homogeneity",
                           "Var: Country Intermediate Class ",
                           "Var: Country Working Class",
                           "Cov: Country (Intercept), Homogeneity",
                           "Cov: Country (Intercept), Intermediate Class ",
                           "Cov: Country (Intercept), Working Class",
                           "Cov: Country Homogeneity, Intermediate Class",
                           "Cov: Country Homogeneity, Working Class",
                           "Cov: Country Intermediate Class, Working Class",
                           "Var: Residual"),
       use.packages = F)
```

@@

st(dfreg,labels=T,
   title="Descriptive Statistics for Study Variables",
   out = "latex",
   digits = 1, 
   factor.counts = F,
   factor.numeric = F,
   summ = list(
     c('notNA(x)','mean(x)','sd(x)','min(x)','max(x)'),
     c('notNA(x)','mean(x)')
   ),
   summ.names = list(
     c('N','Mean','SD','Min','Max'),
     c('Count','Percent')   
   ))
```

@@

\pagebreak 
## Income Inequality Ratios

@@

\pagebreak
## OECD Countries

\pagebreak
## Alternatives

@@
